\chapter{Basic Expansions And Regularization}
\section{Introduction}
\noindent\textbf{Core Idea}: augment/replace the vector of inputs $X$ with additional variables. 

Let $h_m(X):\mathbb{R}^p\rightarrow\mathbb{R}$ be the $m$th transformation, 
\begin{equation*}
    f(X)=\sum_{m=1}^{M} \beta_{m} h_{m}(X)
\end{equation*}
A linear basis expansion in $X$. Once $h_m$ determined, models are linear in these new variables. 
Examples include
\begin{itemize}
\item $h_m(X)=X_m$
\item $H_m(X)=X_j^2$ or $H_m(X)=X_jX_k$. Enable us to augment inputs to achieve higher order 
Taylor expansions. 
\item $H_m(X)=\log(X_j)$, or other nonlinear transformations like $\|X\|$
\item $H_m(X)=I(L_m\le X_j\le U_m$, indicator for a region  
\end{itemize}
Let a dictionary $\mathcal{D}$ consists of all typical basis functions, we have 
\textbf{Three common approaches} to control model complexity
\begin{itemize}
\item Restriction methods, where we decide before-hand to limit the class of functions.
\item Selection methods, which adaptively scan the dictionary and include
only those basis functions $h_m$ that contribute significantly to the fit of
the model.
\item Regularization methods where we use the entire dictionary but restrict
the coefficients.
\end{itemize}

\section{Piecewise Polynomials and Splines}
Assume $X$ is one-dimensional. 
A piecewise polynomial is representing $f$ by different polynomials on different intervals. 
Let $h_m(X)=I(X\in[\xi_{m-1},\xi_{m}])$, $f(X)=\sum_m\beta_m h_m(X)$. Least square estimate
amounts to $\beta_m=\hat{Y}_m$, mean of $Y$ on the $m$th region. 

To make the function continuous, it requires $f(\xi_m^-)=f(\xi_m^+)$. A direct way is using
basis $1,X,(X-\xi_m)_+$. For smoother function, we may increase the order of local polynomials. 

\noindent\textbf{Cubic Spline}: Continuous and first/second derivatives continuous. 

As for order-$M$ spline with $K$ knots $\xi_j$, it should have continuous derivatives up to
order $M-2$. Its base is 
\begin{align*}
h_{j}(X) &=X^{j-1}, j=1, \ldots, M \\ 
h_{M+\ell}(X) &=\left(X-\xi_{\ell}\right)_{+}^{M-1}, \ell=1, \ldots, K
\end{align*}
Cubic spline: lowest-order spline without visible knot-discontinuity. \textbf{Seldom any good 
reason to go beyond that}. 

\subsection{Natural Cubic Splines}
Behavior of polynomial fit to data may be erratic near boundaries, which can be dangerous and
even more wild than global polynomials.  

\textit{Natural Cubic Splines} adds more constraints: function is linear beyond knots. A price
in bias will be paid. 

A natural cubic spline with $K$ knots can be represented by $K$ basis functions. One can start
from a basis for cubic splines and reduce it by boundary constraints. 

\section{Filtering and Feature Extraction}
Preprocess a high dimensional $x$ to some new features by $x^*=g(x)$. 

\section{Smoothing Splines}
Penalized residual sum of squares
\begin{equation}\label{PRSS}
    \operatorname{RSS}(f, \lambda)=\sum_{i=1}^{N}\left\{y_{i}-f\left(x_{i}\right)\right\}^{2}
    +\lambda \int\left\{f^{\prime \prime}(t)\right\}^{2} d t
\end{equation}
$\lambda$: fixed smoothing parameter. Here the first term measures closeness of the data
and the second penalizes curvature. 

Criterion (\ref{PRSS}) is defined on a Sobolev space, which is a space of functions for which the second
term is defined. It can be shown that it has an explicit unique finite-dimensional minimizer
with a natural spline with $N$ knots. Since the solution is a natural spline, we can write it
as
\begin{equation*}
    f(x)=\sum_{j=1}^{N} N_{j}(x) \theta_{j}
\end{equation*}
Here $N_j(x)$ is an $N$-dimensional set of basis for representing this family of natural 
splines. The criterion then reduces to
\begin{equation*}
    \operatorname{RSS}(\theta, \lambda)=(\mathbf{y}-\mathbf{N} \theta)^{T}(\mathbf{y}-\mathbf{N} 
    \theta)+\lambda \theta^{T} \mathbf{\Omega}_{N} \theta
\end{equation*}
where $\{\mathbf{N}_{ij}=N_j(x_i)\}$ and 
$\left\{\Omega_{N}\right\}_{j k}=\int N_{j}^{\prime \prime}(t) N_{k}^{\prime \prime}(t) d t$.
The solution seems to be 
\begin{equation*}
    \hat{\theta}=\left(\mathbf{N}^{T} \mathbf{N}+\lambda 
    \mathbf{\Omega}_{N}\right)^{-1} \mathbf{N}^{T} \mathbf{y}
\end{equation*}
a generalized ridge regression, and fitted spline is given by
\begin{equation*}
    \hat{f}(x)=\sum_{j=1}^{N} N_{j}(x) \hat{\theta}_{j}
\end{equation*}

\subsection{Degrees of Freedom and Smoother Matrices}
Here we discuss intuitive ways of prespecifying $\lambda$. 

A smooth spline with prechosen $\lambda$ is a linear smoother since $\hat{\theta}$ is a linear
combination of $y$. Let $\hat{\mathbf{f}}$ be the estimated values at $x_i$, 
\begin{equation*}
\hat{\mathbf{f}}=\mathbf{N}\left(\mathbf{N}^{T} \mathbf{N}
+\lambda\mathbf{\Omega}_{N}\right)^{-1}\mathbf{N}^{T}\mathbf{y}
=\mathbf{S}_{\lambda} \mathbf{y}
\end{equation*}
$\mathbf{S}_{\lambda}$ known as smoother matrix. 

Let $\mathbf{B}_\xi$ be $N\times M$ matrix of $M$ spline basis functions at $x_i$. $M \ll N$. 
Then least squares fitted spline value is defined by
\begin{align*} 
\hat{\mathbf{f}}&=\mathbf{B}_{\xi}\left(\mathbf{B}_{\xi}^{T} \mathbf{B}_{\xi}\right)^{-1}
\mathbf{B}_{\xi}^{T} \mathbf{y}=\mathbf{H}_{\xi} \mathbf{y} 
\end{align*}
$\mathbf{H}_{\xi}$ is a project operator(hat matrix). $M=tr(\mathbf{H}_\xi)$ gives the 
dimension of the projection space, which is also the number of basis functions/parameters
involved in the fit. By analogy we define \textit{effective degrees of freedom of a smoothing
spline} is
\begin{equation*}
    \mathrm{df}_{\lambda}=\operatorname{trace}\left(\mathbf{S}_{\lambda}\right)
\end{equation*}
We can then find $\lambda$ by specifying $\mathrm{df}_{\lambda}$. 

Since $\mathbf{S}_{\lambda}$ is symmetric, it has a real decomposition, and we can rewrite
it in Reinsch form
\begin{equation*}
    \mathbf{S}_{\lambda}=(\mathbf{I}+\lambda \mathbf{K})^{-1}
\end{equation*}
where $\mathbf{K}$ does not depend on $\lambda$. 
$\hat{\mathbf{f}}=\mathbf{S}_{\lambda} \mathbf{y}$ actually solves
\begin{equation*}
    \min _{\mathbf{f}}(\mathbf{y}-\mathbf{f})^{T}(\mathbf{y}-\mathbf{f})
    +\lambda \mathbf{f}^{T} \mathbf{K} \mathbf{f}
\end{equation*}
$\mathbf{K}$ known as a penalty matrix. Let $d_k$ be the $k$th eigenvalue of $\mathbf{K}$, 
the eigen-decomposition of $\mathbf{S}_{\lambda}$ is
\begin{equation*}
    \mathbf{S}_{\lambda}=\sum_{k=1}^{N} \rho_{k}(\lambda) \mathbf{u}_{k} \mathbf{u}_{k}^{T}
    =\sum_{k=1}^{N} \frac{1}{1+\lambda d_{k}} \mathbf{u}_{k} \mathbf{u}_{k}^{T}
\end{equation*}
By eigen-representation, we can also find that
\begin{itemize}
\item Eigenvectors not affected by changes in $\lambda$. 
\item $\mathbf{S}_{\lambda \mathbf{Y}}=\sum_{k=1}^{N} \mathbf{u}_{k} 
\rho_{k}(\lambda)\left\langle\mathbf{u}_{k}, \mathbf{y}\right\rangle$, hence the smoothing
spline operates by decomposing $\mathbf{y}$ w.r.t $\{\mathbf{u}_k\}$ and differentially 
shrinking the contributions using $\rho_k(\lambda)$, while $\mathbf{H}_{\xi}$ has $M$ 
eigenvalues equal to $1$ and the rest are $0$. So smoothing splines are shrinking smoothers, 
regression splines are projection smoothers. 
\item $\mathbf{u}_k$ ordered by decreasing $\rho_k(\lambda)$ increase in complexity. Higher-complexity
$\mathbf{u}_k$ are shrunk more. If domain of $X$ are periodic, $\mathbf{u}_k$ will be sin/cos. 
\item First two eigenvalues always 1(two-dimensional eigenspace of functions linear in $x$). 
\item $d_1=d_2=0$ and linear functions are not penalized. 
\item Reparametrizing using Demmler-Reinsch basis $\mathbf{u}_k$ solves
\begin{equation*}
    \min _{\boldsymbol{\theta}}\|\mathbf{y}-\mathbf{U} \boldsymbol{\theta}\|^{2}+\lambda 
    \boldsymbol{\theta}^{T} \mathbf{D} \boldsymbol{\theta},\quad \mathbf{D}=diag\{d_k\}
\end{equation*}
\end{itemize}
A smoothing spline is a local fitting method, much like the locally weighted regression. 
As $\lambda\rightarrow 0,~\mathrm{df}_{\lambda} \rightarrow N,~
\mathbf{S}_{\lambda} \rightarrow \mathbf{I}$, As 
$\lambda\rightarrow \infty,~\mathrm{df}_{\lambda} \rightarrow 2,~
\mathbf{S}_{\lambda} \rightarrow \mathbf{H}$, the hat matrix for linear regression. 

\section{Automatic Selection of the Smoothing Parameters}
\subsection{Fixing the Degree of Freedom}
$\mathrm{df}_{\lambda}=\operatorname{trace}\left(\mathbf{S}_{\lambda}\right)$ is monotone in 
$\lambda$. We can set $\mathrm{df}_{\lambda}$ to fix $\lambda$. We may try some different df 
and use some traditional criteria like F-tests, residual plots or others. 

\subsection{The Bias-Variance Tradeoff}
Since $\hat{\mathbf{f}}=\mathbf{S}_{\lambda} \mathbf{y}$, let $\operatorname{Cov}(\mathbf{y})=I$
\begin{align*}
\operatorname{Cov}(\hat{\mathbf{f}})=&\mathbf{S}_{\lambda}\operatorname{Cov}(\mathbf{y})
\mathbf{S}_{\lambda}^{T}=\mathbf{S}_{\lambda} \mathbf{S}_{\lambda}^{T}\\
\operatorname{Bias}(\hat{\mathbf{f}})=&\mathbf{f}-\mathrm{E}(\hat{\mathbf{f}})
=\mathbf{f}-\mathbf{S}_{\lambda} \mathbf{f}
\end{align*}
where $\mathbf{f}$ is the (unknown) vector of evaluations of the true $f$. 

The integrated squared prediction error (EPE) combines both bias and
variance in a single summary: 
\begin{align*} 
\operatorname{EPE}\left(\hat{f}_{\lambda}\right)&=\mathrm{E}\left(Y-\hat{f}_{\lambda}(X)\right)^{2}
=\operatorname{Var}(Y)+\mathrm{E}\left[\operatorname{Bias}^{2}\left(\hat{f}_{\lambda}(X)\right)
+\operatorname{Var}\left(\hat{f}_{\lambda}(X)\right)\right] \\ 
&=\sigma^{2}+\operatorname{MSE}\left(\hat{f}_{\lambda}\right) 
\end{align*}
EPE is a natural quantity of interest, and does create a tradeoff between bias and variance.
When we don’t know the true function, we do not have access to EPE, and need an estimate. Overall
$N-fold$ cross-validation curve could be a good estimation of EPE. 

\section{Nonparametric Logistic Regression}
Consider a general logistic model
\begin{align*}
    \log \frac{\operatorname{Pr}(Y=1 | X=x)}{\operatorname{Pr}(Y=0 | X=x)}&=f(x)\\
    \operatorname{Pr}(Y=1 | X=x)&=\frac{e^{f(x)}}{1+e^{f(x)}}
\end{align*}
To fit $f(x)$ in a smooth fashion, consider the penalized log-likelihood criterion
\begin{align*}
\ell(f ; \lambda) 
&=\sum_{i=1}^{N}\left[y_{i}\log p\left(x_{i}\right)+\left(1-y_{i}\right)\log\left(1-p\left(x_{i}\right)\right)\right]-\frac{1}{2} \lambda \int\left\{f^{\prime \prime}(t)\right\}^{2}dt\\
&=\sum_{i=1}^{N}\left[y_{i} f\left(x_{i}\right)-\log \left(1+e^{f\left(x_{i}\right)}\right)\right]-\frac{1}{2} \lambda \int\left\{f^{\prime \prime}(t)\right\}^{2} dt
\end{align*}
We can also represent $f(x)=\sum_{j=1}^{N} N_{j}(x) \theta_{j}$, and
\begin{align*} 
\frac{\partial \ell(\theta)}{\partial \theta} 
&=\mathrm{N}^{T}(\mathrm{y}-\mathrm{p})-\lambda \mathrm{\Omega} \theta \\ 
\frac{\partial^{2} \ell(\theta)}{\partial \theta \partial \theta^{T}} 
&=-\mathrm{N}^{T} \mathrm{WN}-\lambda \Omega 
\end{align*}
where $\mathrm{W}$ is a diagonal matrix of $p(x_i)(1-p(x_i))$. 
The first derivative is nonlinear in $\theta$, using Newton-Rapthson in method, we can get
update equation
\begin{align*} 
\theta^{\text {new }}
&=\left(\mathbf{N}^{T}\mathbf{W N}+\lambda\mathbf{\Omega}\right)^{-1}\mathbf{N}^{T}
\mathbf{W}\left(\mathbf{N} \theta^{\text {old }}+\mathbf{W}^{-1}(\mathbf{y}-\mathbf{p})\right)
=\left(\mathbf{N}^{T}\mathbf{W}\mathbf{N}+\lambda\mathbf{\Omega}\right)^{-1}\mathbf{N}^{T}\mathbf{W}\mathbf{z} \\
\mathbf{f}^{\text {new }}&=\mathbf{N}\left(\mathbf{N}^{T} \mathbf{W} \mathbf{N}+\lambda \mathbf{\Omega}\right)^{-1} \mathbf{N}^{T} \mathbf{W}\left(\mathbf{f}^{\text {old }}+\mathbf{W}^{-1}(\mathbf{y}-\mathbf{p})\right)
=\mathbf{S}_{\lambda, w}\mathbf{z}
\end{align*}
see that the update fits a weighted smoothing spline to the working response $\mathbf{z}$. 

\section{Multidimensional Splines}
Now we consider multi-dimensional $X$. Let $X\in\mathbb{R}^2$, then the $M_1\times M_2$ dimensional
\textit{tensor product basis} is defined by
\begin{align*}
g_{j k}(X)=&h_{1 j}\left(X_{1}\right) h_{2 k}\left(X_{2}\right), j=1, \ldots, M_{1}, k=1, 
\ldots, M_{2}\\
g(X)=&\sum_{j=1}^{M_{1}} \sum_{k=1}^{M_{2}} \theta_{j k} g_{j k}(X)
\end{align*}
Smoothing spline via regularization can also be constructed by
\begin{align*}
    \min _{f} \sum_{i=1}^{N}\left\{y_{i}-f\left(x_{i}\right)\right\}^{2}+\lambda J[f]
\end{align*}
whose solution has the form
\begin{equation*}
    f(x)=\beta_{0}+\beta^{T} x+\sum_{j=1}^{N} \alpha_{j} h_{j}(x)
\end{equation*}
where $h_{j}(x)=\left\|x-x_{i}\right\|^{2} \log \left\|x-x_{i}\right\|$ are radial basis functions. 

Complexity would be $O(N^3)$. In practice, it is usually sufficient to work with $K$ knots covering
the domain. , which reduces the computations to $O(NK^2+K^3)$. 

Additive spline models are a restricted class of splines where $f(X)=\alpha+\sum f_i(X_i)$, which
might also be more natural, and can be extended to ANOVA spline decompositions. 
\begin{equation*}
    f(X)=\alpha+\sum_{j} f_{j}\left(X_{j}\right)+\sum_{j<k} f_{j k}\left(X_{j}, X_{k}\right)+\cdots
\end{equation*}

Things to consider: 
\begin{itemize}
\item Maximum order of iteraction
\item Terms to include
\item Representation: Groups of small basis versus a large complete basis
\end{itemize}

\section{Regularization and Reproducing Kernel Hilbert Spaces}
Cast splines into the larger context of regularization methods
and reproducing kernel Hilbert spaces. 

A general class has the form
\begin{equation*}
    \min _{f \in \mathcal{H}}\left[\sum_{i=1}^{N} L\left(y_{i}, f\left(x_{i}\right)\right)
    +\lambda J(f)\right]
\end{equation*}
where $\mathcal{H}$ is a space where $J(f)$ is defined. A general penalty functional has the 
form
\begin{equation*}
    J(f)=\int_{\mathbb{R}^{d}} \frac{|\tilde{f}(s)|^{2}}{\tilde{G}(s)} d s
\end{equation*}
where $\tilde{f}$ is a Fourier transform of $f$, and $\tilde{G}$ is some positive function
which falls to zero as $\|s\|\rightarrow\infty$. Under some additional assumptions, the
solutions have the form
\begin{equation*}
    f(X)=\sum_{k=1}^{K} \alpha_{k} \phi_{k}(X)+\sum_{i=1}^{N} \theta_{i} G\left(X-x_{i}\right)
\end{equation*}
where $\phi_k$ span the null space of the penalty function $J$, and $G$ is the inverse Fourier
transform of $\tilde{G}$. 

\subsection{Spaces of Functions Generated by Kernels}
An important subclass of the general regularization problems are generated by positive
definite kernel $K(x,y)$, and the corresponding Hilbert space $\mathcal{H}_K$ is called a \textit{reproducing
kernel Hilbert space}(RKHS). $J$ is defined in terms of kernels as well. 

Let $x,y\in\mathbb{R}^p$, we consider the space of functions generated by the span of 
$\{K(\cdot,y),y\in\mathbb{R}^p\}$. Suppose $K$ has an eigen-expansion
\begin{equation*}
    K(x, y)=\sum_{i=1}^{\infty} \gamma_{i} \phi_{i}(x) \phi_{i}(y),\quad
    \gamma_{i} \geq 0, \sum_{i=1}^{\infty} \gamma_{i}^{2}<\infty
\end{equation*}
Elements of $\mathcal{H}_K$ have an expansion in terms pf eigen functions
\begin{equation*}
f(x)=\sum_{i=1}^{\infty} c_{i} \phi_{i}(x), \quad 
J(f):=\|f\|_{\mathcal{H}_{K}}^{2} \stackrel{\text { def }}{=} \sum_{i=1}^{\infty} c_{i}^{2} 
/ \gamma_{i}<\infty
\end{equation*}
It is a generalized ridge penalty, and functions with larger eigenvalues get penalized less. 

Rewriting the general form we have
\begin{equation*}
\min _{f \in \mathcal{H}_{K}}\left[\sum_{i=1}^{N} L\left(y_{i}, f\left(x_{i}\right)\right)+
\lambda\|f\|_{\mathcal{H}_{K}}^{2}\right]=
\min _{\left\{c_{j}\right\}_{1}^{\infty}}\left[\sum_{i=1}^{N} L\left(y_{i}, \sum_{j=1}^{\infty} c_{j} \phi_{j}
\left(x_{i}\right)\right)+\lambda \sum_{j=1}^{\infty} c_{j}^{2} / \gamma_{j}\right]
\end{equation*}
It can be shown that the solution is finite-dimensional and has the form
\begin{equation*}
    f(x)=\sum_{i=1}^{N} \alpha_{i} K\left(x, x_{i}\right)
\end{equation*}
The basis function $h_i(x)=K(x,x_i)$ is the \textit{representer of evaluation} at $x_i$ in 
$\mathcal{H}_K$, for $f\in\mathcal{H}_K$, $\left\langle K\left(\cdot, x_{i}\right), f\right
\rangle_{\mathcal{H}_{K}}=f\left(x_{i}\right),~\left\langle K\left(\cdot, x_{i}\right), 
K\left(\cdot, x_{j}\right)\right\rangle_{\mathcal{H}_{K}}=K\left(x_i, x_{j}\right)$, hence
\begin{equation*}
    J(f)=\sum_{i=1}^{N} \sum_{j=1}^{N} K\left(x_{i}, x_{j}\right) \alpha_{i} \alpha_{j}
    \text{ for }f(x)=\sum_{i=1}^{N} \alpha_{i} K\left(x, x_{i}\right)
\end{equation*}

The finite-dimensional criterion of the general form is then
\begin{equation*}
    \min _{\alpha} L(\mathbf{y}, \mathbf{K} \boldsymbol{\alpha})+\lambda \boldsymbol{\alpha}^{T}
     \mathbf{K} \boldsymbol{\alpha}
\end{equation*}

\noindent\textbf{Bayesian interpretation}: $f$ is a zero-mean stationary Gaussian with prior
covariance $K$. 

A general case is $\mathcal{H}=\mathcal{H}_0\oplus \mathcal{H}_{1}$ where null space 
$\mathcal{H}_0$ consisting of functions that do not get penalized. The penalty becomes 
$J(f)=\|P_1f\|$ where $P_1$ is the orthogonal projection onto $\mathcal{H}_1$. The solution
then has the form 
$f(x)=\sum_{j=1}^{M} \beta_{j} h_{j}(x)+\sum_{i=1}^{N} \alpha_{i} K\left(x, x_{i}\right)$. 
From a Bayesian perspective, the components in $\mathcal{H}_0$ have improper priors(with
variance). 

\subsection{Examples of RKHS}
\begin{equation*}
    \min _{\alpha}(\mathbf{y}-\mathbf{K} \boldsymbol{\alpha})^{T}(\mathbf{y}-\mathbf{K} 
    \boldsymbol{\alpha})+\lambda \boldsymbol{\alpha}^{T} \mathbf{K} \boldsymbol{\alpha}
\end{equation*}

\subsubsection{Penalized Polynomial Regression}
The kernel $K(x,y)=(\langle x,y\rangle+1)^d$, for $x,y\in\mathbb{R}^p$, has $C_{p+d}^d$
eigen-functions that span the space of $d$-dim polynomials in $\mathbb{R}^p$. 


\subsubsection{Gaussian Radial Basis Functions}
Gaussian kernel $K(x, y)=e^{-\nu\|x-y\|^{2}}$ is chosen because of its functional form in 
the representation $f(x)=\sum_{i=1}^{N} \alpha_{i} K\left(x, x_{i}\right)$. Gaussian radial
basis functions are
\begin{equation*}
    k_{m}(x)=e^{-\nu\left\|x-x_{m}\right\|^{2}}, m=1, \ldots, N
\end{equation*}
The coefficients are estimated with $\min _{\alpha}(\mathbf{y}-\mathbf{K} \boldsymbol{\alpha})^{T}(\mathbf{y}-\mathbf{K} 
\boldsymbol{\alpha})+\lambda \boldsymbol{\alpha}^{T} \mathbf{K} \boldsymbol{\alpha}$. 

\subsubsection{Support Vector Classifiers}
Supporting vector machines 
$f(x)=\alpha_{0}+\sum_{i=1}^{N} \alpha_{i} K\left(x, x_{i}\right)$ where the parameters are
chosen to minimize
\begin{equation*}
    \min _{\alpha_{0}, \alpha}\left\{\sum_{i=1}^{N}\left[1-y_{i} f\left(x_{i}\right)\right]_{+}
    +\frac{\lambda}{2} \alpha^{T} \mathbf{K} \alpha\right\}
\end{equation*}
where $y_i\in\{0,1\}$. It is a quadratic optimization problem with linear constraints. 

\section{Wavelet Smoothing}
Wavelets typically use a complete orthonormal basis to represent functions,
but then shrink and select the coefficients toward a sparse representation.
They are able to represent both
smooth and/or locally bumpy functions in an efficient way—a phenomenon
dubbed time and frequency localization. In contrast, the traditional Fourier
basis allows only frequency localization.

\subsection{Wavelet Bases and the Wavelet Transform}
Bases are generate by translations and dilations of a single scaling function $\phi(x)$(\textit{father}). 
For example, consider Haar basis, space $V_j$ spanned by $\phi_{j.k}=2^{j/2}\phi(2^jx-k)$, 
$\phi(x)=I(x \in[0,1])$. $\cdots \supset V_{1} \supset V_{0} \supset V_{-1} \supset 
\cdots$ ($V_0$ the reference space $V_0$). 

As for the definition of wavelets, we want to represent a function in $V_{j+1}$ by a component in 
$V_j$ plus in the orthogonal complement $W_j$ of $V_j$ to $V_{j+1}$, written as $V_{j+1}=V_{j} \oplus W_{j}$. 
$W_j$ represents detail, and we might wish to set some elements of it to zero. 
Functions $\psi(x-k)$ generated by the mother wavelet $\psi(x)=\phi(2x)-\phi(2x-1)$ form an orthonormal
basis for $W_0$ for the Haar family, and $\psi_{j, k}=2^{j / 2} \psi\left(2^{j} x-k\right)$ form
a basis for $W_j$. 
We also know $V_{J}=V_{0} \oplus W_{0} \oplus W_{1} \cdots \oplus W_{J-1}$. 

Since the spaces are orthogonal, all basis functions should be orthonormal. 

\subsubsection{Tradeoff between Wavelet Basis}
\begin{itemize}
\item The wider the support, the more time the wavelet has to die to zero, so it can achieve this more
smoothly. Note that the effective support seems to be much narrower.
\item $V_0$ is equivalent to the null space of the smoothing-spline penalty. 
\end{itemize}

\subsection{Adaptive Wavelet Filtering}
Wavelets are very useful when the data are measured on a uniform lattice(discretized signal, 
image, or time series). Here we consider one-dimensional case with $N=2^J$ lattice-points. 
Let $\mathbf{W}$ be the $N\times N$ orthonormal wavelet basis matrix evaluated at the $N$ 
uniformly spaced observations, $\mathbf{y}$ be the response vector, Then $\mathbf{y}^*=
\mathbf{W}\mathbf{y}$ is the \textit{wavelet transform} of $\mathbf{y}$. A popular method for 
adaptive wavelet fitting is SURE shrinkage. We start with
\begin{equation*}
    \min _{\boldsymbol{\theta}}\|\mathbf{y}-\mathbf{W} \boldsymbol{\theta}\|_{2}^{2}+2 
    \lambda\|\boldsymbol{\theta}\|_{1}
\end{equation*}
which is same as the lasso criterion. Since $\mathbf{W}$ is orthonormal, 
\begin{equation*}
    \hat{\theta}_{j}=\operatorname{sign}\left(y_{j}^{*}\right)
    \left(\left|y_{j}^{*}\right|-\lambda\right)_{+}
\end{equation*}
The fitted function (vector) is then given by the inverse wavelet transform $\hat{\mathbf{f}}^*=
\mathbf{W}\hat{\mathbf{y}}$. 

A simple choice for $\lambda$ is $\lambda=\sigma \sqrt{2 \log N}$($\sigma$ the estimate of the 
standard deviation). The reason is when $\mathbf{y}$ are white noise, so are $\mathbf{y}^*$. 
This $\lambda$ is the expected maximum of $N$ white noise with variance $\sigma^2$. 

% TODO: Appendix on computations for splines